Personalization and probabilities: impersonal propensities in online grocery shopping

### 2017-09-08 log.rmd
29/31 pages edited
31/31 pages entered

## Thu 31 Aug 2017 16:34:43 BST
- long gap here -- due to important other activities such as summer holidays;  
- found the 1989 Humby article on demographics -- argues for the need to add lifestage to that of neighbourhood. 
- TODO: add reference to [@Humby_1989]

>>>>

## Thu 31 Aug 2017 16:34:43 BST
- long gap here -- due to important other activities such as summer holidays;  
- found the 1989 Humby article on demographics -- argues for the need to add lifestage to that of neighbourhood. 
- DONE: add reference to [@Humby_1989]
- reached conclusion in adding edits. Not sure if the section on Popper and possibiities trying to realize themselves is fleshed out enough.
- TODO: check if the last section does enough with propensities and the new features.

## Fri 08 Sep 2017 09:24:22 BST
- re-writing the conclusion today - did this on paper and now in file; it still needs more work
- but in order to do that, I've had to go back to beginning and go through the paper looking at the argument again.< 
- TODO: jump between D. and E. needs fixing. 
- TODO: section on list theory is still awkward -- Miller is fine, but what then?

### 2017-09-08 reconstruction.rmd
31/31 pages entered

---
title: "Personalization and probabilities: impersonal propensities in online grocery shopping"
author: Adrian Mackenzie^[Sociology, Lancaster University]
date: "`r format(Sys.time(), '%B %Y')`"
fontsize: 12pt
citation_package: biber
latex_engine: xelatex
bibliography: /home/mackenza/ref_bibs/uni.bib
csl: "sage-harvard.csl"
link-citations: yes
header-includes:
    - \usepackage{setspace}
    - \doublespacing
abstract:
    "Many accounts of big data platforms assume that they target individuals. Personalization, with all the risks of discrimination and bias it entails, has been the focus of inquiry. This paper suggests that personalization occurs against a more expansive background that in principle is not reducible to individual attributes and desires, but includes many different forms of relations in flux. It describes the 'personalization' of an online grocery shopping recommender system. It reconstructs an account of how a recommender system constructs an ordered list of a small number of grocery items of personal relevance for each of the millions of online grocery shoppers in  a major UK supermarket chain. Drawing on a theory of probability proposed by the philsopher of science Karl Popper, it suggests that personal relevance is inevitably bound up with relations to others, and the relations between things. Using a mixture of discourse analysis and code-based reconstruction of key elements of the recommender system, it suggests that personalization is one facet of a much more open-ended weave of propensities associated with people and things in contemporary big data configurations.   The paper explores how in the context of recommender systems the constitutive incompleteness of shopping lists, their propensity to expand or change, might be more important than their capacity to be personalized."
---

@@

This paper reconstructs [@Dewey_1957] an online grocery ordering system undergoing an upgrade from a demographically-based recommendation system to a 'personal relevance' model. The reconstruction is empirical in certain respects. It departs from an ethnographic moment: being a member of the audience at an industry/academic conference where a data scientist was describing how recommendations were created for online grocery orders. Starting from this highly mediated and fragmentary encounter with a recommender system and one of its designers, the paper explores the problem of personally relevant recommendations.  It draws on anthropological and sociological research concerning shopping and list-making to situate the recommender system amongst the social ordering practices of shopping and lists. It makes use of archaeological approaches, in the sense developed by Michel Foucault [@Foucault_1972] in order to identify and map important functional statements and practices configuring the knowledge and order generated by such systems over time. It conducts several small-scale code-based experiments in order to reconstruct, using widely available code resources resources such as API (Application Programmer Interfaces) and software libraries for machine learning, some prototypical elements of the system in question.  Throughout, it draws  on the philosophy of science and media theory to frame an engagement with data, algorithms and recommendations in terms of probabilities and their worldly enactments.

The main argument of the paper concerns how  personalization occurs in Big Data situations. In recent research and opinion on data practice, personalization has been seen as the linchpin of predictive modelling.  Big data discourse in its promissory mode  attributes potency to personalization: 'most important, using big data we hope to identify specific individuals rather than groups' [@Mayer-Schonberger_2013]. Conversely, Joseph Turow and co-authors conclude in their critical account of the transformation of retail space: 'through it all, knowingly and not, and away from the spotlights of fierce social debate, retailers are encouraging daily routines that accept data-driven personalization as a centrifugal public' [@Turow_2015, 476]. Analysis of internet filter bubbles and the growth of predictive platforms in industry and government make similar arguments [@Pariser_2011; @ONeil_2016a].

@@

Against the background of the sheer number of commodities and their distribution of prices,  Patel's presentation presents two opportunities. First, by recounting, contextualising and commenting on the main steps in making the 'Have your forgotten?' list, we might follow some of the predictive sense-making done by data scientists and customer analytics teams working with transactional data in a typical commercial setting. Patel mentioned many of these steps only fleetingly in the presentation, for they are largely taken for granted as part of predictive analytic practice. Second, Patel focused on the renovation and updating of long-standing data-mining practice via a much more explicitly 'big data' and 'machine learning'- oriented implementation. Her presentation concerns what we might term, for reasons that will gradually emerge, a 'conversion event' in which a long-standing recommender system was replaced by a 'big data'-style system concerned with 'personally relevant' recommendations.   What stands out from the presentation is the normality of the recommender system: it is part of a long-standing and ongoing transformation or conversion of grocery shopping. From a sociological standpoint, the interest lies less in specific technical innovations and more on how the ordering work done by recommender systems relates to the role of prediction in social order more generally.

@@

Online grocery shopping at [Tesco](https://www.tesco.com/groceries/?icid=dchp_groceriesshopgroceries), the largest supermarket chain in the UK, includes recommendations for further grocery purchases under the title of 'Have you forgotten?'. When Tesco customers shop for groceries online, a list of five recommendations appear at the checkout stage. The recommendations are the product of a recommender system, an important category of operational device in big data (see for instance [@Hallinan_2014] for analysis of Netflix recommendation; [@Morris_2015]  or [@Seaver_2015] for an account of music recommendations).  The question 'have you forgotten?' is followed by a list of some  grocery items that could have been or are usually on a shopping list. The title of the suggestions is a bit misleading: the recommender system, as we will see, is not concerned with forgetting, with the many slips and oversights associated to which shopping is prone,  but rather with items that customers had not selected perhaps because they had never thought of buying them in the first place.

The anthropologist  Daniel Miller has argued [@Miller_2012] that  shopping takes place as a negotiation of discrepancies between normative and actual social order (for instance, between the ideals of health emblematised by organic products and commitments to thrift embodied in generic products). More provocatively, Miller claims that 'shopping is largely a technology for the expression of love' (85). What Miller suggests here is that shopping practices attempt to resolve difference between how people think they should live and how they actually live: 'we have to watch how shopping helps resolve these discrepancies between the normative and the actual, but we also need some ideas of where the normative comes from in the first place' [@Miller_2012, 72]. Whatever theory we have of how normative social order arises, the key argument is that shopping inhabits a space between normative/idea and actual order. Importantly, and this will be a key consideration for recommender systems, shopping is not particularly personal or individual. Grocery shopping in particular is saturated  by fluxing forms of social order concerning family and other forms of social grouping, and their associated relations (love, etc.).

Shopping lists provide important clues to how local social order is constituted, maintained, and repaired in shopping. If social ordering is done in shopping,  then shopping list as an inscriptive device stabilises certain forms of ordering at the expense of other possible orderings. The shopping list, whether written on the back of an envelope, or saved as a list in an online grocery shopping system, lies at the intersection of logistical flows, infrastructural orderings, and  lively negotiations around actual and normative social orders.[^30.1]  Shopping lists, therefore  are intersectional ordering devices that encapsulates a universe of possible references, and a teeming multitude of propensities with an actual local order (in both senses of that term: an ordering and an imperative). As people shop, either by  trawling along aisles packed with thousands of products, or scrolling down screens or searching for particular brands amidst search results, lists inscribe some kind of order that filters or reduces the excessive propensities, claims, dazzle and distraction of commodities to the practicalities of domestic economy.

@@

Hand-written shopping lists undoubtedly have ongoing importance and display interestingly mixed ordering practices (see the montage of handwritten shopping lists at [Grocery List](http://www.grocerylist.org)). But online shopping lists take shape at the intersection of web and internet infrastructures, supply chain logistics, individualised practices and *habitus,* and increasingly, the predictive operations of  recommender systems. Just as the aisles and shelves of a supermarket present a densely-woven semiotics of objects competing for visual attention by offering distinctions of taste, thrift, expedience, novelty, indulgence, health and increasingly, online shopping recommender systems generate lists that seek to align people to commodities that they otherwise might have little relation to (see [@Turow_2015] for an overview).

@@

The main narrative of Patel's presentation concerned a shift from a well-established loyalty-card based data-mining model to a predictive, probabilistic, personally relevant re-writing of the shopping list in almost-realtime. The changes Patel described are increasingly widespread and common. While they are configured in Tesco-specific ways by DunnHumby (and this reflects a longer history), they are also broadly typical. Tesco is the largest supermarket chain in UK, famous amongst retailers for its  customer loyalty and targeted marketing programme known as 'Tesco Clubcard,' which started in 1991. DunnHumby -- founded by operations researchers Edwina Dunn and Clive Humby -- is said to have convinced the CEO of Tesco sometime in 1991 that a loyalty card program could change the supermarket chain's relationship to its customers.  Clive Humby's academic publications are hard to track down (see [@Humby_1989] where Humby argues for the need to add lifestage data to neighbourhood data in marketing research), but an early paper given at the _Conference of Young Operational Researchers_ in Nottingham in 1984 suggests the direction that he, DunnHumby and later Tesco would take in constructing lists [@OKeefe_1984]. The abstract for Humby's presentation pre-figures an ongoing trajectory for data mining techniques aimed at eliciting detailed information on individual customer references.

@@

While Tesco  succeeded in data-mining its customers using demographic segmentation,  and perhaps became the UK's biggest supermarket with the help of data-mining in the 1990s, the shopping environment  in 2017 is markedly different [@Turow_2017]. It is no longer organised around campaigns involving special offers or redemption of points for demographically segmented loyal customers (DunnHumby made heavy use of UK Census data). It can no longer rely on placement of goods in carefully chosen locations in stores. It needs or at least might want to continuously up-sell and cross-sell to customers who might hardly ever visits the supermarket itself.

How could we characterise the shift from ClubCard data mining to online grocery markets?  If 'Tesco is the clear winner in the online grocery market, in fact it takes almost 50p of every £1 spent on food shopping on the internet' [How SEO helps Tesco to dominate the online grocery market](https://econsultancy.com/blog/64841-how-seo-helps-tesco-to-dominate-the-online-grocery-market)[@Silverwood-Cope_2014], then Tesco itself has undergone some kind of conversion? Patel described her work at DunnHumby as shifting the recommender system from a 'rules-based list' to a 'relevancy model.' The relevancy model changed how the 'Have You Forgotten' list was constructed.   This shift is a typical of broader re-organisation of prediction in which existing data analysis practices are being re-distributed and intensified in particular ways. The shift in models results in a more probabilistic structuring of lists.  The ostensible banality of the 'Have you forgotten?' recommendations betrays micrological signs of a topologically complex predictive infrastructure that intervenes in list-based social ordering to reconfigure alignments between people and things.

This is not a drastic or abrupt event. Academic researchers first began writing about recommender systems in the mid-1990s. From the outset they highlighted a potential shift from demographic-driven market research or data mining techniques to personalized recommendations. For instance, writing in 1997 in a special section of the _Communications of the ACM_ on recommender systems, Paul Resnick and Hal Varian (at that time Dean of Information Sciences at UC Berkeley, but currently Chief Economist at Google), made much of this personalisation [@Resnick_1997]. Resnick and Varian  emphasise the need to distinguish the emerging practices from data mining:

> In everyday life, we rely on recommendations from other people. ... Recommender systems augment this natural social process. In a typical recommender  system, people provide recommendations as inputs, which the system then aggregates and directs to appropriate recipients. In some cases the primary transformation is in the aggregation; in others the system’s value lies in its ability to make good matches between the recommenders and those seeking recommendations.[@Resnick_1997,56]

Writing just after the advent of web-based e-Commerce, Resnick and Varian  imagined recommender systems augmenting the 'collaborative filtering' that people do when they write and read reviews of products and services (for instance, on the travel website, TripAdvisor; [@Scott_2012]). In 1997,  Resnick and Varian expected that 'people [would] provide recommendations' and recommender systems would aggregate and rank recommendations for recipients. A slightly later review, [@Schafer_2001], diagrams an augmented 'natural social process' with a range of elements, technologies, inputs and outputs, with varying degrees of personalization.

@@

The 'minute and extensive scrutinizing' [@Dewey_1957] of a recommender system undergoing a conversion from an older data-mining approach to a current 'personal relevance model' approach addresses a specific situation: online grocery shopping (not shopping in general) and, even more micro-sociologically, the ordering of lists.

The theory of shopping lists constructed at the beginning of the paper was a scaffolding component. It draws on anthropology, sociology and mathematics to suggest that we see shopping lists as the site of a complex negotiation between different social orders, between norms and actualities. The list itself as an inscriptive device functions as an ordering device, a device that accommodates an immense variety of specific habits, dispositions, values and associations. The mapping element of the list, the order that arranges the set of items according to different interests, values, and aims, is the operational focus of recommender systems.

The personalization of recommendations has been a defining characteristic of many big data conversion events. Rightly, the effects of all-consuming personalization have been the target of critics and critical research. The archaeology of the Tesco recommender system, including its early experiments with demographic data-mining, its adoption of a rule-based 'market basket' analysis system, and its current implementation of a 'personal relevance model' certainly follows that trajectory.   But the conversion event is not reducible to personalization.

Probabilization precedes personalization. When we observe what is practically done to construct a personal relevance model and apply it to a shopping list, we see, conditional probabilities, predictive models, the importance of data accumulated over time,  platform experiences, the rhythms of infrastructural maintenance, and open-ended construction of new features in data, all of which negotiate shifting and complicated associations between people and things, people and people, and things with things. If social order is made of associations, is to be social is to associate, then the forms of association taking shape in this and other similar conversion events might be understood as understand as operationalizing probability in matrices of propensity. As Popper argues, probabilities are being realized in the same way that fields of force were realized in modern physics. Probabilization does not take place in physics laboratories but on platforms, combining databases, compute capacity, interfaces and global supply chain logistics.

The fact that probabilisation largely takes the form of personalization tends to obscure the re-framing of social order taking place. This ordering  depends on  the mappings that inscriptive devices such as shopping lists  display. Only because shopping is already somewhat ordered by lists can a partial re-mapping, provisionally conducive to new events, take place. The conversion entailed in this re-mapping is not just about personalization. It is more ontological than that since it includes relations running multi-laterally between people and things in time. It is less ontological, however, in that it is not a sudden transformation, but an ongoing process of adjusting, scaffolding, intervening and configuring orders.

>>>>

---
title: "Personalization and probabilities: impersonal propensities in online grocery shopping"
author: Adrian Mackenzie^[Sociology, Lancaster University]
date: "`r format(Sys.time(), '%B %Y')`"
fontsize: 12pt
citation_package: biber
latex_engine: xelatex
bibliography: /home/mackenza/ref_bibs/uni.bib
csl: "sage-harvard.csl"
link-citations: yes
header-includes:
    - \usepackage{setspace}
    - \doublespacing
abstract:
    "Many accounts of big data platforms assume that they target individuals. Personalization, with all the risks of discrimination and bias it entails, has been the focus of inquiry. This paper suggests that personalization occurs against a more expansive background of probabilization that in principle is not reducible to individual attributes and desires, but includes many different forms of relations in flux. It describes the 'personalization' of an online grocery shopping recommender system. It reconstructs an account of how a recommender system arranges an ordered list of a small number of grocery items of personal relevance for each of the millions of online grocery shoppers in  a major UK supermarket chain. Drawing on a theory of probability proposed by the philsopher of science Karl Popper, it suggests that personal relevance is inevitably caught up in relations to others, and the relations between things. Using a mixture of discourse analysis and code-based reconstruction of key elements of the recommender system, it suggests that personalization is one facet of a much more open-ended weave of propensities associated with people and things in contemporary big data configurations. The paper explores how in the context of recommender systems the constitutive incompleteness of shopping lists, their propensity to expand or change, might be more important than their capacity to be personalized."
---

@@

## Introduction

This paper reconstructs [@Dewey_1957] an online grocery ordering system undergoing an upgrade from a demographically-based recommendation system to a 'personal relevance' model. The reconstruction is empirical in certain respects. It departs from an ethnographic moment: being a member of the audience at an industry/academic conference where a data scientist was describing how personalized recommendations were created for online grocery orders. Starting from this highly mediated and fragmentary encounter with a recommender system and one of its designers, the paper explores the problem of personally relevant recommendations.  It draws on anthropological and sociological research concerning shopping and list-making to situate the recommender system amongst the social ordering practices of shopping and lists. It makes use of archaeological approaches, in the sense developed by Michel Foucault [@Foucault_1972] in order to identify and map important functional statements and practices configuring the knowledge and order generated by such systems over time. It conducts several small-scale code-based experiments in order to reconstruct, using widely available code resources resources such as API (Application Programmer Interfaces) and software libraries for machine learning, some prototypical elements of the system in question.  Throughout, it draws  on the philosophy of science and media theory to frame an engagement with data, algorithms and recommendations in terms of probabilities and their worldly enactments.

The main argument of the paper concerns how  personalization occurs in Big Data 'conversion events' (a term that refers to the common before-after narrative forms that proponents, developers and advocates of these systems adopt). In recent research and opinion on data practice, personalization has been seen as the linchpin of predictive modelling.  Big data discourse in its promissory mode  attributes potency to personalization: 'most important, using big data we hope to identify specific individuals rather than groups' [@Mayer-Schonberger_2013]. Conversely, Joseph Turow and co-authors conclude in their critical account of the transformation of retail space: 'through it all, knowingly and not, and away from the spotlights of fierce social debate, retailers are encouraging daily routines that accept data-driven personalization as a centrifugal public' [@Turow_2015, 476]. Analysis of internet filter bubbles and the growth of predictive platforms in industry and government make similar arguments [@Pariser_2011; @ONeil_2016a].

@@

Against the background of the sheer number of commodities and their distribution of prices,  Patel's presentation presents two opportunities. First, by recounting, contextualising and commenting on the main steps in making the 'Have your forgotten?' list, we might follow some of the predictive sense-making done by data scientists and customer analytics teams working with transactional data in a typical commercial setting. Patel mentioned many of these steps only fleetingly in the presentation, for they are largely taken for granted as part of predictive analytic practice. Second, Patel focused on the renovation and updating of long-standing data-mining practice via a much more explicitly 'big data' and 'machine learning'- oriented implementation. Her presentation concerns what we might term, for reasons that will gradually emerge, a 'conversion event' in which a long-standing recommender system was replaced by a 'big data'-style system concerned with 'personally relevant' recommendations.   What stands out from the presentation is the normality of the recommender system: it is part of a long-standing and ongoing transformation of grocery shopping. From a sociological standpoint, the interest lies less in specific technical innovations and more on how the ordering work done by recommender systems relates to the role of prediction in social order more generally.

@@

Online grocery shopping at [Tesco](https://www.tesco.com/groceries/?icid=dchp_groceriesshopgroceries), the largest supermarket chain in the UK, includes recommendations for further grocery purchases under the title of 'Have you forgotten?'. When Tesco customers shop for groceries online, a list of five recommendations appear at the checkout stage. The recommendations are the product of a recommender system, an important category of operational device in big data (see for instance [@Hallinan_2014] for analysis of Netflix recommendation; [@Morris_2015]  or [@Seaver_2015] for an account of music recommendations).  The question 'have you forgotten?' is followed by a list of some  grocery items that could have been or are usually on a shopping list. The title of the suggestions is a bit misleading. The recommender system, as we will see, is not concerned with forgetting, with the many slips and oversights associated to which shopping is prone,  but rather with items that customers had not selected perhaps because they had never thought of buying them in the first place.

The anthropologist Daniel Miller has argued [@Miller_2012] that  shopping takes place as a negotiation of discrepancies between normative and actual social order (for instance, between the ideals of health emblematised by organic products and commitments to thrift embodied in lower cost generic products). What Miller suggests here is that shopping practices attempt to resolve difference between how people think they should live and how they actually live: 'we have to watch how shopping helps resolve these discrepancies between the normative and the actual, but we also need some ideas of where the normative comes from in the first place' [@Miller_2012, 72]. More provocatively, Miller claims that 'shopping is largely a technology for the expression of love' (85). It is, to use the language of actor-network theory, a practice of translation [@Callon_1991]. Whatever theory we have of how normative social order arises, the key implication is that shopping inhabits a mundane space between normative/idea and actual order. Importantly, and this will be a key consideration for recommender systems, shopping is not particularly personal or individual. Grocery shopping in particular is saturated  by fluxing forms of social order concerning family and other forms of social grouping, and their associated relations (love, etc.).

If social ordering is done in shopping, then what do shopping lists do?  Shopping lists provide important clues to how local social order is constituted, maintained, and repaired in shopping. As an inscriptive device, a list stabilises certain quite complex forms of social order at the expense of other possible orderings.  As people shop, either by  trawling along aisles packed with thousands of products, or scrolling down screens or searching for particular brands amidst search results, lists inscribe some kind of order that filters or reduces the excessive propensities, claims, dazzle and distraction of commodities to the practicalities of domestic economy. The shopping list, whether written on the back of an envelope, saved as a list in an online grocery shopping system, or remembered, lies at the intersection of logistical flows, infrastructural orderings, and  lively negotiations around actual and normative social orders.[^30.1]  Shopping lists  are intersectional ordering devices that encapsulate a universe of possible references, and a teeming multitude of propensities with an actual local order.

@@

Hand-written shopping lists undoubtedly have ongoing importance and display interestingly mixed ordering practices (see the montage of handwritten shopping lists at [Grocery List](http://www.grocerylist.org)). But online shopping lists take shape at the intersection of web and internet infrastructures, supply chain logistics, individualised practices and *habitus,* and increasingly, the predictive operations of  recommender systems. Just as the aisles and shelves of a supermarket present a densely-woven semiotics of objects competing for visual attention by offering distinctions of taste, thrift, expedience, novelty, indulgence, health and increasingly, online shopping recommender systems generate lists that seek to align people to commodities that they otherwise might have little relation to (see [@Turow_2015] for an overview of the development of these systems).

@@

The main narrative of Patel's conversion narrative concerned the shift from a well-established loyalty-card based data-mining model to a predictive, probabilistic, personally relevant construction of the shopping list in almost-realtime. While the changes Patel described are configured in Tesco-specific ways by DunnHumby, they are also broadly typical. Tesco is the largest supermarket chain in UK, famous amongst retailers for its  customer loyalty and targeted marketing programme known as 'Tesco Clubcard,' which started in 1991. DunnHumby -- founded by operations researchers Edwina Dunn and Clive Humby -- is said to have convinced the CEO of Tesco sometime in 1991 that a loyalty card program could change the supermarket chain's relationship to its customers.  Clive Humby's academic publications are hard to track down (see [@Humby_1989] where Humby argues for the need to add lifestage data to neighbourhood data in marketing research), but an early paper given at the _Conference of Young Operational Researchers_ in Nottingham in 1984 suggests the direction that he, the company Dunn and Humby formed (DunnHumby) and later Tesco would take in constructing lists [@OKeefe_1984]. The abstract for Humby's presentation pre-figures an ongoing trajectory for data mining techniques aimed at eliciting detailed information on individual customer references.

@@

While Tesco  succeeded in data-mining its customers using demographic segmentation,  and perhaps became the UK's biggest supermarket with the help of data-mining in the 1990s, the shopping environment  in 2017 is markedly different [@Turow_2017]. It is no longer organised around campaigns involving special offers or redemption of points for demographically segmented loyal customers (DunnHumby made heavy use of UK Census data). It can no longer rely on placement of goods in carefully chosen locations in stores. It needs or at least might want to continuously up-sell and cross-sell to customers who might hardly ever visit the supermarket itself.

How could we characterise the shift from ClubCard data mining to online grocery markets?  If 'Tesco is the clear winner in the online grocery market, in fact it takes almost 50p of every £1 spent on food shopping on the internet' [How SEO helps Tesco to dominate the online grocery market](https://econsultancy.com/blog/64841-how-seo-helps-tesco-to-dominate-the-online-grocery-market)[@Silverwood-Cope_2014], then has Tesco itself undergone some kind of conversion? Patel described her work at DunnHumby as converting the recommender system from a 'rules-based list' to a 'relevancy model.' The relevancy model changed how the 'Have You Forgotten' list was constructed.   This shift is a typical of broader re-organisation of prediction in which existing data analysis practices are being re-distributed and intensified in particular ways. The shift in models results in a more probabilistic structuring of lists.  The ostensible banality of the 'Have you forgotten?' recommendations betrays micrological signs of a topologically complex predictive infrastructure that intervenes in list-based social ordering to reconfigure alignments between people and things.

This is not a drastic or abrupt event. Academic researchers first began writing about personalized recommender systems in the mid-1990s. From the outset they highlighted a potential shift from demographic-driven market research or data mining techniques to personalized recommendations. For instance, writing in 1997 in a special section of the _Communications of the ACM_ on recommender systems, Paul Resnick and Hal Varian (at that time Dean of Information Sciences at UC Berkeley, but currently Chief Economist at Google), made much of this personalization [@Resnick_1997]. Resnick and Varian  emphasise the need to distinguish the emerging practices from data mining:

>In everyday life, we rely on recommendations from other people. ... Recommender systems augment this natural social process. In a typical recommender  system, people provide recommendations as inputs, which the system then aggregates and directs to appropriate recipients. In some cases the primary transformation is in the aggregation; in others the system’s value lies in its ability to make good matches between the recommenders and those seeking recommendations.[@Resnick_1997,56]

Writing just after the advent of web-based e-Commerce, Resnick and Varian  imagined recommender systems drawing on  the 'collaborative filtering' that people do when they write and read reviews of products and services (for instance, on the travel website, TripAdvisor; [@Scott_2012]). In 1997,  Resnick and Varian expected that 'people [would] provide recommendations' and recommender systems would aggregate and rank recommendations for recipients. A slightly later review, [@Schafer_2001], diagrams an augmented 'natural social process' with a range of elements, technologies, inputs and outputs, with varying degrees of personalization.

@@

The personalization of recommendations has been a defining characteristic of many big data conversion events. The exploitative and discriminatory effects of all-consuming personalization have been the target of critics and critical research on data economies. The archaeology of the Tesco recommender system, including its early experiments with demographic data-mining, its adoption of a rule-based 'market basket' analysis system, and its current implementation of a 'personal relevance model' follows the trajectory of personalization.

The fact that the conversion event largely takes the form of personalization,  an individualizing personalization that is so patently ill-fit to the socially complex negotiations of grocery shopping, tends to obscure the wider re-framing of social order taking place.  on  the mappings that inscriptive devices such as shopping lists  display. Only because shopping is already somewhat ordered by lists can a partial re-mapping, provisionally conducive to new events, take place. The conversion entailed in this re-mapping is not just about personalization. It is more ontological than that since it includes relations running multi-laterally between people and things in time. It is less ontological, however, in that it is not a sudden transformation, but an ongoing process of adjusting, scaffolding, intervening and configuring orders.

But the conversion event associated with big data is not reducible to personalization or the individual. Probabilization precedes and exceeds personalization. When we observe what is practically done to construct a personal relevance model and apply it to a shopping list, we see conditional probabilities, predictive models, the importance of data accumulated over time,  platform experiences, the rhythms of infrastructural maintenance, and open-ended construction of new features in data, all of which shift and complicate the conditional associations between people and things, people and people, and things with things. If social order is made of associations, if to be social is to associate [@Latour_TBA], then the associations taking shape in this and other similar big data conversion events might be understood as operationalizing probability in matrices of propensity. If, as Popper argues, probabilities are being realized in the same way that 'fields' were realized in modern physics, then we should see the compute  platforms, databases, software libraries, web interfaces and global supply chain logistics as part of this conversion event.
